Endogenous Politics and \\ the Design of Trade Institutions

### 2017-01-01 presentation.tex
move repeated game intuition slide to appendix
reverse order of conclusion slides
move lit review to obj fcn section
first compile of presentation for ASSAs, recompile to-do

\title[Endogenous Politics and the Design of Trade Institutions\hspace{2.05in}\insertframenumber/\inserttotalframenumber]{Endogenous Politics and \\ the Design of Trade Institutions}
%\author[Kristy Buzard]{Kristy Buzard \\ Syracuse University and The Wallis Institute \\ kbuzard@syr.edu}
\author[Kristy Buzard]{\texorpdfstring{Kristy Buzard\newline Syracuse University and The Wallis Institute  \newline\url{kbuzard@syr.edu}}{Kristy Buzard}}
\date{June 30, 2016}
\begin{document}
\maketitle
%\insertpresentationendpage removed b/c of appendix

@@

\section{Objective Fcn}
\subsection{Objective Function}
\begin{frame}{Restraining Political Pressure through TAs}
\pause
%It's easy to see that weak bindings keep lobbying alive relative to strong bindings
\begin{itemize}[<+->]
	\item Will TA be used to discourage lobbying? Depends on how gov't welfare varies in $\ga$
	\item With standard Baldwin-style objective function, welfare always increases with $\ga$\spaceskip1pt
\[
  W = \mathit{CS}_X(\tau) + \ga \pi_X(\tau) + \mathit{CS}_Y(\tau^*) + \pi_Y(\tau^*) + \mathit{TR}(\tau)
\]

	\begin{itemize}
		\item Note `Protection for Sale' objective function is isomorphic
	\end{itemize}

@@

\begin{frame}{Tariff Caps with Self Enforcement}
\pause
\begin{itemize}[<+->]
	\item $\ga$ exogenous (Bagwell $\&$ Staiger 2005): if governments patient enough ($\de$ high enough), optimal externally-enforced weak binding can be self-enforced
  \item $\ga$ endogenous: optimal externally-enforced weak binding may not be self-enforcing
		\begin{itemize}
			\item Problem: lobby is an additional repeated-game player
			\item Lobby's incentive constraint is harder to satisfy as $\de$ increases
		\end{itemize}

\end{itemize}

@@

\begin{frame}{Repeated Game Intuition}
\pause
Legislature: break agreement if punishment not strong enough
\pause
\begin{itemize}
	\item i.e. if one period of gain from cheater's payoff is greater than T-periods of loss from trade-war
\end{itemize}

\pause
\vskip.1in
Executives: set lowest $\bta$ that makes paying $\ov{e}_b$ unprofitable \textit{and} satisfies legislature's condition
\pause
\begin{itemize}
	\item[$\Rightarrow$] $e_b=0$, agreement remains in force
	\pause
	\item High tariffs, no lobbying, no trade disruptions
\end{itemize}
\end{frame}

@@

@@

\begin{frame}{Conclusion}
Taking into account endogenous political forces alongside exogenous ones...
\pause
\begin{itemize}[<+->]
		\item provides additional general explanation for tariff caps
		\item demonstrates that TAs can be used to discourage lobbing activity in general
		\item helps explain the structure and enforcement of the WTO Safeguards measure
\end{itemize}

>>>>

\usepackage{appendixnumberbeamer}

\title[Endogenous Politics and the Design of Trade Institutions\hspace{2.05in}\insertframenumber/\inserttotalframenumber]{Endogenous Politics and \\ the Design of Trade Institutions}
%\author[Kristy Buzard]{Kristy Buzard \\ Syracuse University and The Wallis Institute \\ kbuzard@syr.edu}
\author[Kristy Buzard]{\texorpdfstring{Kristy Buzard\newline Syracuse University \newline\url{kbuzard@syr.edu}}{Kristy Buzard}}
\date{January 6, 2017}
\begin{document}
\maketitle
%\insertpresentationendpage removed b/c of appendix

@@

\section{Objective Fcn}

\subsection{Objective Function}
\begin{frame}{Restraining Political Pressure through TAs}
\pause
%It's easy to see that weak bindings keep lobbying alive relative to strong bindings
\begin{itemize}[<+->]
	\item Will TA be used to discourage lobbying? Depends on how gov't welfare varies in $\ga$
	\item With standard Baldwin-style objective function, welfare always increases with $\ga$\spaceskip1pt
\[
  W = \mathit{CS}_X(\tau) + \ga \pi_X(\tau) + \mathit{CS}_Y(\tau^*) + \pi_Y(\tau^*) + \mathit{TR}(\tau)
\]

	\begin{itemize}
		\item Isomorphic to `Protection for Sale' objective function
	\end{itemize}

@@

\begin{frame}[label=self]
\frametitle{Tariff Caps with Self Enforcement}
\pause
\begin{itemize}[<+->]
	\item $\ga$ exogenous (Bagwell $\&$ Staiger 2005): if governments patient enough ($\de$ high enough), optimal externally-enforced weak binding can be self-enforced
  \item $\ga$ endogenous: optimal externally-enforced weak binding may not be self-enforcing
		\begin{itemize}
			\item Problem: lobby is an additional repeated-game player
			\item Lobby's incentive constraint is harder to satisfy as $\de$ increases
		\end{itemize}

\end{itemize}
\pause
\hyperlink{repeated<1>}{\beamergotobutton{Repeated Game Intuition}}
\end{frame}

@@

@@

\begin{frame}{Conclusion}
Taking into account endogenous political forces alongside exogenous ones...
\pause
\begin{itemize}[<+->]
		\item helps explain the structure and enforcement of the WTO Safeguards measure
		\item can help us think about optimal design of trading insitutions
		\item demonstrates that TAs can be used to discourage lobbing activity in general
		\item provides additional general explanation for tariff caps
\end{itemize}

@@

\appendix
\begin{frame}[label=repeated]
\frametitle{Repeated Game Intuition}
\pause
Legislature: break agreement if punishment not strong enough
\pause
\begin{itemize}
	\item i.e. if one period of gain from cheater's payoff is greater than T-periods of loss from trade-war
\end{itemize}

\pause
\vskip.1in
Executives: set lowest $\bta$ that makes paying $\ov{e}_b$ unprofitable \textit{and} satisfies legislature's condition
\pause
\begin{itemize}
	\item[$\Rightarrow$] $e_b=0$, agreement remains in force
	\pause
	\item High tariffs, no lobbying, no trade disruptions
\end{itemize}
\hyperlink{self<1>}{\beamergotobutton{Go Back}}
\end{frame}
